---
title: "ESM 206 Lab 6"
author: "Tommy King"
date: "11/2/2021"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired or independent 2-sample t-test). 

H0: Equal ranks (or medians)
HA: Ranks are not equal (or medians are not equal)

### Create samples

`set.seed()` to create "pseudorandom but reproducible sample"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424) 
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributinos using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so that we can compare the results. 
```{r}
ttest <- t.test(gp_1, gp_2)

# if these data were paired, I can just add the argument `paired = "TRUE" ` within the t-test() function.
```

This p-value means that if our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is `r round(ttest$p.value, 2)` probability of finding sample means *at least as different* as mine by random chance. 

[A more comprehensive explanation of diffrernce between groups... actually means, data spread, effect size]... there is no significant difference (t(df) = t-statistic value, p =).


### Mann Whitney U rank based test

We will use the `wilcox.test()` function to performa a two-sample rank-based test. 

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that if our samples are drawn from populations with the same median, there is a 28% chance of finding sample medians at least as different as ours by random chance. 

NO signigicant difference in medians (Mann Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value,2)`)

Exmaple text for reporting: "The medians of group A (M = 22.4g) and group B (M= 31.2 g) did not differ significantly (Mann Whitney U, ...)" 

## SImple Linear regression 

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think tthat a linear relationship makes sense. 

Use the `penguins` data from {palmerpenguins}, exploring flipper length (x) versus body mass (y). 

```{r}
# MAke an exploratory scatterplot of penguin flipper length on x-axis and body mass on y-axis. 
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_jitter(aes(color = species)) + 
  geom_smooth(method = "lm")
```


### Find the Model estimates or coefficients 

Use the `lm()` function for linear model


```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

```

Model Coefficients: `r penguin_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients[2]`

### USe `broom::tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy 
```

### Use `broom::glance()` to get output of model fit and statistical significance 

```{r}
penguin_lm_glance <- glance(penguin_lm)

```

### Use equatiomatic:: extract_eq() to return model in LaTex
```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE) # add the `use_coefs()` function to put in the actual coefficient values in the equation

```

### USe plot() function to automatically produce diagnostic plots

```{r}
plot(penguin_lm)

```

Residuals: difference or variance between actual values and what my model predicts them to be. (Line of best fit vs. actual plot points)
