---
title: "ESM 206 Lab 6"
author: "Tommy King"
date: "11/2/2021"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired or independent 2-sample t-test). 

H0: Equal ranks (or medians)
HA: Ranks are not equal (or medians are not equal)

### Create samples

`set.seed()` to create "pseudorandom but reproducible sample"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424) 
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributinos using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so that we can compare the results. 
```{r}
ttest <- t.test(gp_1, gp_2)

# if these data were paired, I can just add the argument `paired = "TRUE" ` within the t-test() function.
```

This p-value means that if our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is `r round(ttest$p.value, 2)` probability of finding sample means *at least as different* as mine by random chance. 

[A more comprehensive explanation of diffrernce between groups... actually means, data spread, effect size]... there is no significant difference (t(df) = t-statistic value, p =).


### Mann Whitney U rank based test

We will use the `wilcox.test()` function to performa a two-sample rank-based test. 

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that if our samples are drawn from populations with the same median, there is a 28% chance of finding sample medians at least as different as ours by random chance. 

NO signigicant difference in medians (Mann Whitney U; U = `r mwu$statistic`, p = `mwu$p.value,2)`)
